---
title: "Oil Accidents Analysis"
author: "Janick Reales & Luisa Rios"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# OIL PIPELINE ACCIDENTS

The data used in this analysis was taken from [kaggle](https://www.kaggle.com/usdot/pipeline-accidents). The goal of the analysis is to undertand the main aspects relaten with the collected data. 


Firt, we have to import the libraries needed.
```{r echo=FALSE}
library(tidyverse)
library(shiny)
library(dplyr)
library(ggplot2)
library(plotly)
library(leaflet)
library(htmltools)
library(RColorBrewer)
library(kableExtra)
```

#### Importing database
```{r echo=FALSE}
oil_accidents <- read.csv("database.csv", header=TRUE, sep=',', encoding="UTF-8")
```


The imported database contains `r dim(oil_accidents)[1]` rows and `r dim(oil_accidents)[2]` columns. The following table shows the fisrt rows of the dataset. 

```{r}
dt <- head(oil_accidents)
 kable(dt) %>%
      kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F) %>%
      scroll_box(width = "910px", height = "500px")
```

Where the number of Na/Null values and the type of attribute in each column are:
```{r}
## counting the num of NA values in each column
na_values <- colSums(is.na(oil_accidents))

## checking the type of attribute in each column
col_type <- sapply(oil_accidents, class)


###
type_values <- data.frame(na_values, col_type)

 kable(type_values) %>%
      kable_styling(bootstrap_options = c("hover", "condensed"), full_width = T) %>%
      scroll_box(width = "400px", height = "500px")
```

Ana the 
```{r}

```

